---
title: "GOLDHR"
author: "GroupVAHAC"
date: "1/26/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo=FALSE, message=FALSE}
source("library.R")
```

```{r echo=FALSE, message=FALSE}
suppressWarnings(source("library.R"))
# Package options
suppressWarnings(ggthemr('fresh'))  # ggplot theme
opts_knit$set(progress=FALSE, verbose=FALSE)
opts_chunk$set(echo=FALSE, fig.align="center", fig.width=10, fig.height=6.35, results="asis")
options(knitr.kable.NA = '')
 
Datafile = "data/HR_comma_sep.csv"
```
 
```{r}
ProjectData <- read.csv(Datafile)
ProjectData <- data.matrix(ProjectData)
ProjectData_INITIAL <- ProjectData
```

## Objectives of the project


## Data set
Load the data from Kaggle.com regarding HR analytics in a major US company

https://www.kaggle.com/ludobenistant/hr-analytics

```{r echo=FALSE}
local_directory <- getwd()
ProjectData <- read.csv(file = "data/HR_comma_sep.csv", header = TRUE, sep=",")
```

Let's visualize the sample data for 10 employees

```{r echo=FALSE}
# let's visualize the sample data for 10 employees

ProjectData1 <- ProjectData[,1:18]

ProjectData1 = data.matrix(ProjectData1)
```

```{r echo=FALSE}
max_data_report <- 10

knitr::kable(round(head(ProjectData1, max_data_report), 2))
```

## Descriptive statistics

We show the descriptive statistics of the factors

```{r echo=FALSE}
my_summary <- function(thedata){
  res = apply(thedata, 2, function(r) c(min(r), quantile(r, 0.25), quantile(r, 0.5), mean(r), quantile(r, 0.75), max(r), sd(r)))
  res <- round(res,2)
  colnames(res) <- colnames(thedata)
  rownames(res) <- c("min", "25 percent", "median", "mean", "75 percent", "max", "std")
  t(res)
}

ProjectData1 <- ProjectData[,1:18]

knitr::kable(round(my_summary(ProjectData1), 2))

```

We need to scale the data (excluding dummy variables for the department)

```{r, echo=FALSE, tidy=TRUE}
ProjectData2 <- ProjectData[,1:9]
ProjectData2_scaled=apply(ProjectData2,2, function(r) {if (sd(r)!=0) res=(r-mean(r))/sd(r) else res=0*r; res})
```

Notice now the summary statistics of the scaled dataset (excluding dummy variables for the department):

```{r echo=FALSE}
ProjectData2 <- ProjectData[,1:9]
knitr::kable(round(my_summary(ProjectData2_scaled),2))
```

Let's see how these are correlated. The correlation matrix is as follows:

```{r}
show_data = round(cor(ProjectData1),2)

knitr::kable(show_data)
```

## Dimensionability reduction *(second priority)*

## Cluster Analysis

## Predictive machine learning (artificial stupidity)

## Executive summary and final conclusions

## Recommendations for refinement

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
