---
title: "GOLDHR"
author: "GroupVAHAC"
date: "1/26/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo=FALSE, message=FALSE}
source("../INSEADAnalytics/AnalyticsLibraries/library.R")
```


## Objectives of the project


## Data set
Load the data from Kaggle.com regarding HR analytics in a major US company

https://www.kaggle.com/ludobenistant/hr-analytics

```{r echo=FALSE eval=TRUE}
local_directory <- getwd()
ProjectData <- read.csv(file = "data/HR_comma_sep.csv", header = TRUE, sep=",")
```

## Descritptive statistics

```{r}
my_summary <- function(thedata){
  res = apply(thedata, 2, function(r) c(min(r), quantile(r, 0.25), quantile(r, 0.5), mean(r), quantile(r, 0.75), max(r), sd(r)))
  res <- round(res,2)
  colnames(res) <- colnames(thedata)
  rownames(res) <- c("min", "25 percent", "median", "mean", "75 percent", "max", "std")
  t(res)
}

ProjectData1 <- ProjectData[,1:6]

knitr::kable(round(my_summary(ProjectData1), 2))
```

```{r}
renderFormattable(dformat(head(ProjectData, SAMPLE_ROWS)))

```


## Dimensionability reduction *(second priority)*

## Cluster Analysis

## Predictive machine learning (artificial stupidity)

## Executive summary and final conclusions

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
